---
title: "Fundamentals"
output: html_document
date: '2022-10-03'
---

# load Tidyverse package and load ITE fake data
```{r}
library(tidyverse)

# Load ITE fake data
load("ite_ml/ite_fake.rda")

```
tips: to run the above code chunk, you can either click on the `play` button on the right or use keyboard shortcut `ctrl+shft+enter` to run the enter code chunk or `ctrl+enter` to run code line by line. Give it a try!

# View ITE fake data
```{r}
# to view it on a separate tab
ite %>% view()

# to view it on console
ite %>% glimpse()
```
Note the symbol %>% means pipe. It takes the dataframe `ite` and piped it to the function `view`

# How to use Select
```{r}
# To select column/variable PGY1_percentile only 
ite %>%
  select(PGY1_percentile)

# To select column/variable PGY1_percentile & PGY1_score only 
ite %>%
  select(PGY1_percentile, PGY1_score)

# To select column/variable PGY1_percentile & PGY1_score only, and assign to a new dataframe called new_df
new_df <- ite %>%
  select(PGY1_percentile, PGY1_score)
```

# How to use Filter
```{r}
# To filter column/variable PGY1_percentile of less than or equal to 35
ite %>%
  filter(PGY1_percentile <= 35)

# To filter column/variable PGY1_percentile of less than or equal to 35 OR PGY1_score less than or equal to 60
ite %>%
  filter(PGY1_percentile <= 35 | PGY1_score <= 60)
```

# How to use group_by and mutate
```{r}
# To group by residents who passed/failed and to add a new column to assess count of Pass=Yes and Pass=No
ite %>%
  group_by(Pass) %>%
  mutate(count = n())
```

# How to use case_when
```{r}
# how to divide PGY1_percentile to low, medium, high 
ite %>%
  mutate(PGY1_percentile_level = case_when(
    PGY1_percentile <= 35 ~ "low",
    PGY1_percentile > 35 && PGY1_percentile <= 60 ~ "medium",
    PGY1_percentile > 60 ~ "high"
  )) %>%
  select(Resident, PGY1_percentile, PGY1_percentile_level) %>%
  arrange(PGY1_percentile)

```

